\large{Vipassana for Hackers} \\ \Huge{The Proposal} \\ \large\textit{Version 0.1}

### 2019-11-29 the-proposal/the-proposal.tex
started intro draft for proposal

\begin{abstract}
Vipassana meditation (as taught by S.N. Goenka in the tradition of Sayagyi U Ba Khin)
is unlike other meditation techniques insofar as its claims to completeness and
outcomes are concerned. Vipassana claims to ultimately explore the entire field of
mind and matter, with the goals of total liberation and full enlightenment. Implicit
within these claims is a complete understanding of human consciousness. These are
difficult claims to prove because the time commitment required to research the
technique orders on multiple decades --- if not multiple generations. This is complicated by the
fact that the time commitment is demanded of both the researcher \textit{and the
subject}. Until now, research on meditation of all kinds has covered only one of
two fields: (1) controlled experiments which must inherently rely on superficial data
gathered from beginners and (2) observational exploratory research of monks --- expert
meditators who have dedicated their lives to the practice. I propose bridging this
gap by submitting myself to the middle ground. While remaining a layperson, I will
commit to a high ratio of waking medition hours for the rest of my life. Internally,
I will conduct qualitative research into the nature of consciousness while
simultaneously exposing myself as a subject to long-term quantitative studies.

@@

\keywords{neuroscience, vipassana, meditation}

@@

% \listoftodos

\section{Why Vipassana?}

zig zig zig

@@

Thank you to Steven Deobald for reviews, edits, and corrections. \texttt{Brain\_human\_normal\_inferior\_view.svg} and
\texttt{Brain\_human\_lateral\_view.svg} by Patrick J. Lynch, medical illustrator, CC-BY 2.5 \cite{brainsvg}

@@

>>>>

\begin{abstract}
Vipassana meditation (as taught by S.N. Goenka in the tradition of Sayagyi U Ba Khin)
is unlike other meditation techniques insofar as its claims to completeness and
outcomes are concerned. Vipassana claims to ultimately explore the entire field of
mind and matter, with the goals of total liberation and full enlightenment. Implicit
within these claims is a complete understanding of human consciousness. These are
difficult claims to prove because the time commitment required to research the
technique orders on multiple decades --- if not multiple generations. This is complicated by the
fact that the time commitment is demanded of both the researcher \textit{and the
subject}. Until now, research on meditation of all kinds has covered only one of
two fields: (1) controlled experiments which must inherently rely on superficial data
gathered from beginners and (2) observational exploratory research of monks --- expert
meditators who have dedicated their lives to the practice. I propose bridging this
gap by submitting myself to the middle ground. While remaining a layperson, I will
commit to a high ratio of waking medition hours for the rest of my life. Internally,
I will conduct qualitative research into the nature of consciousness while
externally exposing myself as a subject to long-term quantitative studies.

@@

\keywords{neuroscience, psychology, vipassana, meditation}

@@

\listoftodos

\section{Introduction}

Research into the effects of meditation has been conducted in earnest for half a
century but the quality of research in this field varies wildly. Randomized
controlled trials were missing from much research conducted during the first few
decades of meditation study. The importance of active controls was often missed even
when randomized controlled trials were attempted. Double-blind studies are inherently
impossible with meditation research; a subject will always know if she is receiving
meditation instruction or an active control instruction. \cite{alteredtraits}

Add to these difficulties the very nature of meditation research itself. There are
many techniques of meditation and it is very important to capture the specific
technique under study to make meaningful assertions about its
effects. \cite{alteredtraits} However, even within the definition of a single
meditation technique there exist variations in instruction between
teachers and each student's comprehension of the instructions
received. \cite{davidson2015} Even if researchers could cement (or at least
accurately record) semantics and terminology, the subject of study is often unclear:
Are we looking for health benefits? Increases in productivity? Increases in
intelligence? How long these effects persist? How much do we want to explore concrete
hypotheses versus exploratory analysis of long-term effects? How much can be learned
about the nature of consciousness? Can these learnings be measured objectively or
communicated meaningfully?

Because Vipassana is globally standardized, it affords researchers with definitive
solutions to the difficulties presented by varying teaching methodologies. In
exchange for this, the difficulty of long-term study is compounded by the intrinsic
seriousness demanded of a Vipassana practitioner: the concept of ``continuous
piercing awareness of constantly changing bodily sensation'' is not only a
requirement of serious Vipassana practice but could very well act as a surrogate
definition for the practice itself. This seriousness poses obvious difficulties.

\subsection{Why Vipassana?}

1. ``complete'', standardized, global, multilingual, 100\% free
2. mundane (sleep) vs. supramundane (total eradication of suffering)

3. who is steven deobald

\section{Problem Statement}

A large scale controlled study of beginner-to-intermediate
students of Vipassana, each practicing a minimum of two hours daily, is
possible. Randomization will not be possible in such a study as the students
self-select this

- middle ground: me

(vs. monks)

\subsection{Overview}
\subsection{Hypothesis}

@@

Thank you to Preethi Govindarajan for reviews, edits, and corrections.

@@

\bibitem{alteredtraits}
  Goleman D., Davidson R.J.
  \textit{Altered Traits.}
  ISBN: 9780399184390. September 2018.

\bibitem{davidson2015}
  Davidson R.J., Kaszniak A.W.
  \textit{Conceptual and methodological issues in research on mindfulness and meditation.}
  American Psychologist, Volume 70(7), Pages 581-92. October 2015.
  \url{https://doi.org/10.1037/a0039512}
